\hebrewsection{Advanced Spectral Analysis}

\subsection{The Discrete Fourier Transform (DFT)}
The DFT $X[k]$ is a sampled version of the Discrete-Time Fourier Transform (DTFT). It represents the signal in the discrete-frequency domain.
\begin{equation}
    X[k] = \sum_{n=0}^{N-1} x[n] e^{-j\frac{2\pi}{N}nk}
\end{equation}

\subsection{Properties of the DFT}
1. \textbf{Linearity:} $DFT\{ax + by\} = aX + bY$.
2. \textbf{Periodicity:} $X[k+N] = X[k]$.
3. \textbf{Circular Shift:} $x[(n-m) \pmod N] \leftrightarrow X[k]e^{-j\frac{2\pi}{N}mk}$.
4. \textbf{Symmetry:} For real signals, $X[k] = X^*[N-k]$.
5. \textbf{Parseval's Relation:} $\sum |x[n]|^2 = \frac{1}{N} \sum |X[k]|^2$.

\subsection{The Fast Fourier Transform (FFT)}
The FFT reduces complexity from $O(N^2)$ to $O(N \log N)$. 

\subsubsection{Radix-2 DIT (Decimation-in-Time)}
The most common algorithm, it recursively breaks an $N$-point DFT into two $N/2$-point DFTs using the "butterfly" structure.
\begin{equation}
    X[k] = G[k] + W_N^k H[k]
\end{equation}

\subsubsection{Radix-4 and Split-Radix Algorithms}
For $N$ being a power of 4, Radix-4 algorithms further reduce the number of multiplications. Split-radix combines the best of both worlds, offering even higher efficiency for specific $N$.

\subsection{Spectral Estimation: Non-Parametric Methods}
\subsubsection{The Periodogram}
The periodogram is a basic estimate of the Power Spectral Density (PSD). However, it is an \textbf{inconsistent estimator} because its variance does not decrease as $N \to \infty$.

\subsubsection{Welch's Method}
Welch's method improves the periodogram by:
1. Dividing the signal into overlapping segments.
2. Windowing each segment.
3. Averaging the periodograms of the segments.
This reduces variance at the expense of frequency resolution.

\subsection{Windowing Functions}
Windows are used to mitigate \textbf{spectral leakage} caused by the finite observation interval (rectangular windowing).
\begin{fancytable}{l c c}
Window & Sidelobe Level (dB) & Main-lobe Width (bins) \\
Rectangular & -13 & 2 \\
Hamming & -43 & 4 \\
Blackman & -58 & 6 \\
Kaiser ($\beta=6$) & -50 & 5 \\
\end{fancytable}

\begin{center}
\begin{tikzpicture}
\begin{axis}[
    title={Spectral Leakage: Rectangular vs. Hamming},
    xlabel={Frequency (Hz)},
    ylabel={Magnitude (dB)},
    grid=both,
    ymin=-60, ymax=0,
    width=0.8\textwidth,
    height=7cm
]
\addplot[blue, domain=-2:2, samples=100] {20*log10(abs(sin(deg(x*pi*5))/(5*sin(deg(x*pi)))))};
\addplot[red, thick, domain=-2:2, samples=100] {20*log10(abs(0.54*sin(deg(x*pi*5))/(5*sin(deg(x*pi))) + 0.23*sin(deg((x-0.2)*pi*5))/(5*sin(deg((x-0.2)*pi))) + 0.23*sin(deg((x+0.2)*pi*5))/(5*sin(deg((x+0.2)*pi)))))};
\end{axis}
\end{tikzpicture}
\end{center}

\subsection{Parametric Spectral Estimation}
Parametric methods assume a model for the signal generation process.
\subsubsection{Autoregressive (AR) Models}
The signal is modeled as the output of an all-pole filter driven by white noise. The parameters are estimated using the \textbf{Yule-Walker equations} or the \textbf{Burg method}. AR models are excellent for resolving closely spaced peaks (spectral lines) in short data records.

\subsubsection{Moving Average (MA) and ARMA Models}
MA models are all-zero, while ARMA models have both poles and zeros. They are used for signals with broader spectral features.

\subsection{High-Resolution Subspace Methods}
\textbf{MUSIC} (Multiple Signal Classification) and \textbf{ESPRIT} exploit the eigen-structure of the autocorrelation matrix. They provide resolution far beyond the traditional DFT limit by separating the signal and noise subspaces.

\subsection{Time-Frequency Analysis: The STFT}
For non-stationary signals, the \textbf{Short-Time Fourier Transform (STFT)} provides a local frequency representation by sliding a window over the signal.
\begin{equation}
    STFT\{x[n]\}(m, \omega) = \sum_{n=-\infty}^{\infty} x[n]w[n-m]e^{-j\omega n}
\end{equation}

\subsubsection{The Spectrogram}
The Spectrogram is the squared magnitude of the STFT. It is a tool for visualizing non-stationary signals like speech and bird songs. The \textbf{Uncertainty Principle} dictates the trade-off between time and frequency resolution.

\subsection{The Discrete Cosine Transform (DCT)}
The DCT uses only cosine basis functions. It has exceptional \textbf{energy compaction} properties, making it the standard for lossy compression (JPEG, MP3).
