\hebrewsection{Digital Filter Design and Implementation}

Digital filters are fundamental components in DSP systems, used to alter the frequency spectrum of signals. This chapter provides a rigorous treatment of both Finite Impulse Response (FIR) and Infinite Impulse Response (IIR) filter design methodologies.

\subsection{Finite Impulse Response (FIR) Filters}
FIR filters are characterized by an impulse response $h[n]$ of finite duration. Their transfer function is a polynomial in $z^{-1}$:
\begin{equation}
    H(z) = \sum_{n=0}^{N-1} h[n] z^{-n}
\end{equation}

\subsubsection{Linear Phase Property}
A key advantage of FIR filters is the ability to achieve exact linear phase, which is impossible for stable IIR filters. Linear phase ensures that all frequency components are delayed by the same amount, preserving the signal's wave shape.
The condition for linear phase is symmetry or anti-symmetry:
\begin{equation}
    h[n] = \pm h[N-1-n]
\end{equation}

\subsubsection{Window Method Design}
The window method starts with the ideal (infinite) impulse response $h_d[n]$ and truncates it using a window function $w[n]$:
\begin{equation}
    h[n] = h_d[n] \cdot w[n]
\end{equation}
Common windows include:
\begin{itemize}
    \item \textbf{Rectangular:} $w[n] = 1$ for $0 \leq n \leq N-1$. High resolution but poor sidelobe rejection (-13 dB).
    \item \textbf{Hamming:} $w[n] = 0.54 - 0.46 \cos\left(\frac{2\pi n}{N-1}\right)$. Good sidelobe rejection (-43 dB).
    \item \textbf{Kaiser:} Allows for a trade-off between main-lobe width and sidelobe level using a parameter $\beta$.
\end{itemize}

\subsubsection{Optimal FIR Design: Parks-McClellan Algorithm}
The Parks-McClellan algorithm, based on the Remez exchange algorithm, produces "equiripple" filters. It minimizes the maximum error in the passband and stopband:
\begin{equation}
    \min_{h[n]} \max_{\omega \in S} W(\omega) |H_d(e^{j\omega}) - H(e^{j\omega})|
\end{equation}
where $W(\omega)$ is a weighting function.

\begin{center}
\begin{tikzpicture}
\begin{axis}[
    title={Equiripple FIR Filter Response},
    xlabel={Normalized Frequency ($\omega/\pi$)},
    ylabel={Magnitude (dB)},
    grid=major,
    ymin=-60, ymax=5,
    width=0.8\textwidth,
    height=7cm
]
\addplot[blue, thick, domain=0:0.4, samples=50] {0.5*cos(deg(x*pi*20))};
\addplot[blue, thick, domain=0.4:0.5, samples=20] {-150*(x-0.4) + 0.5};
\addplot[blue, thick, domain=0.5:1, samples=100] {-50 + 2*sin(deg(x*pi*40))};
\end{axis}
\end{tikzpicture}
\end{center}

\subsection{Infinite Impulse Response (IIR) Filters}
IIR filters use feedback, resulting in an impulse response of infinite duration. They can achieve much sharper transitions than FIR filters for the same order.

\subsubsection{Classic Analog Approximations}
IIR design usually begins by designing an analog prototype and then transforming it to the digital domain.
\begin{enumerate}
    \item \textbf{Butterworth:} Maximally flat magnitude response in the passband.
    \item \textbf{Chebyshev Type I:} Equiripple in the passband, monotonic in the stopband.
    \item \textbf{Chebyshev Type II:} Monotonic in the passband, equiripple in the stopband.
    \item \textbf{Elliptic (Cauer):} Equiripple in both passband and stopband, providing the narrowest transition band for a given order.
\end{enumerate}

\subsubsection{The Bilinear Transform}
The Bilinear Transform is the most common method for mapping the $s$-plane to the $z$-plane. It avoids aliasing by warping the frequency axis:
\begin{equation}
    s = \frac{2}{T} \frac{1 - z^{-1}}{1 + z^{-1}}
\end{equation}
The relationship between analog frequency $\Omega$ and digital frequency $\omega$ is:
\begin{equation}
    \Omega = \frac{2}{T} \tan(\omega/2)
\end{equation}

\subsection{Filter Structures and Finite Word-Length Effects}
Implementing filters in hardware or fixed-point DSPs introduces quantization errors.
\begin{itemize}
    \item \textbf{Coefficient Quantization:} Can move poles outside the unit circle, causing instability.
    \item \textbf{Arithmetic Round-off Noise:} Occurs at the output of multipliers.
    \item \textbf{Limit Cycles:} Oscillations that can occur in IIR filters even with zero input.
\end{itemize}

\begin{pythonbox}{IIR Filter Design via Bilinear Transform}
\begin{verbatim}
import scipy.signal as signal
import numpy as np

# Design a 4th order Chebyshev Type I lowpass filter
# Passband ripple: 1dB, Cutoff frequency: 0.3*fs/2
b, a = signal.cheby1(4, 1, 0.3, btype='low', analog=False)

# Frequency response
w, h = signal.freqz(b, a)
\end{verbatim}
\end{pythonbox}

\subsection{Frequency Sampling Method}
The frequency sampling method specifies the desired frequency response $H(e^{j\omega})$ at $N$ equally spaced points and solves for the impulse response coefficients using the Inverse DFT:
\begin{equation}
    h[n] = \frac{1}{N} \sum_{k=0}^{N-1} H[k] e^{j\frac{2\pi}{N}nk}
\end{equation}
Transition band samples are often optimized to reduce ripple in the resulting continuous response.

\subsection{Comparison of FIR and IIR Filters}
\begin{fancytable}{l c c}
Feature & FIR & IIR \\
Stability & Always Stable & Potentially Unstable \\
Phase & Linear Phase Possible & Non-linear Phase \\
Efficiency & Low (higher order) & High (lower order) \\
Implementation & Non-recursive & Recursive \\
\end{fancytable}
