\hebrewsection{Radar Signal Processing}

Digital Signal Processing is fundamental to modern radar systems. This chapter explores the principles of pulse compression, Doppler processing, and the ambiguity function.

\subsection{Basic Radar Principles}
A radar system transmits a signal $s(t)$ and listens for echoes. The time delay $\Delta t$ between transmission and reception determines the range $R$:
\begin{equation}
    R = \frac{c \Delta t}{2}
\end{equation}
where $c$ is the speed of light.

\subsection{Pulse Compression}
To achieve high range resolution without sacrificing transmit energy, radar systems use long, modulated pulses rather than short, high-power pulses.

\subsubsection{Linear Frequency Modulation (LFM) - Chirp}
The LFM signal, or chirp, increases its frequency linearly over the pulse duration $T$:
\begin{equation}
    s(t) = e^{j\pi K t^2}, \quad -T/2 \leq t \leq T/2
\end{equation}
where $K$ is the chirp rate. At the receiver, a matched filter (pulse compressor) collapses the pulse to a width of approximately $1/B$, where $B=KT$ is the bandwidth.

\subsection{The Ambiguity Function}
The ambiguity function $|\chi(\tau, f_d)|$ represents the output of the matched filter when the received signal has a delay $\tau$ and a Doppler shift $f_d$. It defines the resolution and clutter rejection capabilities of a specific waveform.

\subsection{Doppler Processing and MTI}
Moving Target Indication (MTI) uses the Doppler shift $f_d = \frac{2v}{\lambda}$ to distinguish moving targets from stationary ground clutter. Modern systems use \textbf{Pulse Doppler Radar} with FFT-based processing to resolve targets in both range and velocity.

\begin{pythonbox}{LFM Chirp Generation}
\begin{verbatim}
import numpy as np
import matplotlib.pyplot as plt

T = 10e-6    # 10 microseconds
B = 5e6      # 5 MHz
fs = 50e6    # 50 MHz
t = np.arange(-T/2, T/2, 1/fs)
K = B/T
chirp = np.exp(1j * np.pi * K * t**2)
\end{verbatim}
\end{pythonbox}
