\hebrewsection{Digital Image Processing}

Image processing is the application of DSP techniques to two-dimensional signals. This chapter covers the foundations of image representation, filtering, and compression.

\subsection{2D Signals and Systems}
An image $f(x, y)$ is a 2D signal where $(x, y)$ are spatial coordinates and the amplitude $f$ is the intensity or gray level.

\subsubsection{2D Sampling and Quantization}
Images are sampled on a 2D grid. The resolution depends on the sampling density. Quantization (typically 8-bit for grayscale) determines the number of intensity levels.

\subsection{Image Enhancement in the Spatial Domain}
Spatial domain techniques operate directly on the pixels.
\begin{itemize}
    \item \textbf{Point Processing:} Gamma correction, histogram equalization.
    \item \textbf{Spatial Filtering:} Convolution with kernels for smoothing (Gaussian) or sharpening (Laplacian).
\end{itemize}

\subsection{Filtering in the Frequency Domain}
The 2D DFT allows for filtering in the frequency domain.
\begin{equation}
    F(u, v) = \sum_{x=0}^{M-1} \sum_{y=0}^{N-1} f(x, y) e^{-j2\pi(\frac{ux}{M} + \frac{vy}{N})}
\end{equation}
Low-pass filters blur the image, while high-pass filters enhance edges.

\subsection{Image Compression}
Compression reduces the amount of data required to represent an image.
\begin{enumerate}
    \item \textbf{Lossless (e.g., PNG):} Uses Huffman or LZ coding.
    \item \textbf{Lossy (e.g., JPEG):} Uses the Discrete Cosine Transform (DCT) to remove high-frequency information that is less visible to the human eye.
\end{enumerate}

\subsubsection{The JPEG Standard}
JPEG works by:
1. Dividing the image into 8x8 blocks.
2. Applying DCT to each block.
3. Quantizing the coefficients using a psychovisual matrix.
4. Huffman coding the result.

\subsection{Edge Detection}
Edges are significant local changes in intensity.
\begin{itemize}
    \item \textbf{Sobel Operator:} Approximates the gradient using $3 \times 3$ kernels.
    \item \textbf{Canny Edge Detector:} A multi-stage algorithm that includes noise reduction, gradient calculation, non-maximum suppression, and hysteresis thresholding.
\end{itemize}

\subsection{Wavelet Transforms}
Unlike the Fourier transform, wavelets provide both time (space) and frequency localization. The \textbf{Discrete Wavelet Transform (DWT)} is the basis for JPEG 2000, offering superior compression at low bitrates without the "blocking" artifacts of standard JPEG.
