\hebrewsection{Ethical and Social Implications}

As communication systems become ubiquitous, they bring significant ethical and social challenges that engineers must address.

\subsection{Privacy and Surveillance}
High-resolution radar and Wi-Fi sensing can "see" through walls and track people's movements without their consent. The same DSP algorithms that enable 5G can be used for mass surveillance.

\subsection{Spectrum Equity}
Spectrum is a finite natural resource. How it is allocated (licensed vs. unlicensed) determines who has access to high-speed communication and at what cost. "Spectrum hunger" can lead to the marginalization of rural or low-income communities.

\subsection{Environmental Impact}
The proliferation of electronic devices (IoT) leads to significant e-waste. Furthermore, the massive data centers and cellular networks required for global connectivity consume vast amounts of energy. Engineers must prioritize energy-efficient "green" communications.

\subsection{Security and Resilience}
As society depends on communication for critical infrastructure (power grids, autonomous vehicles), the resilience of these systems against cyberattacks and physical interference is paramount. DSP techniques like \textbf{Physical Layer Security} aim to provide encryption-like protection using the inherent properties of the channel.

\subsection{The Digital Divide}
Despite the advances in 5G and satellite internet, billions of people still lack basic connectivity. Bridging this gap is not just a technical challenge but a social and economic necessity.
